### 2015-06-22 chapters/ipa_meets_unicode.tex
updated handling of arrows, and changed codepoints
see wikipedia proposals: https://en.wikipedia.org/wiki/Upstep

As explained in Section~\ref{pitfall-blocks}, the Unicode Standard code space is
subdivided into character blocks, which generally encode characters from a
single script. However, as is illustrated by the IPA, characters that form a
single writing system may be dispersed across several different character
blocks. With its diverse collection of symbols from various scripts and
diacritics, the IPA is spread across 12 blocks in the Unicode
Standard:\footnote{This number of blocks depends on whether only IPA-sanctioned
symbols are counted or if the phonetic symbols commonly found in the literature
are also included, see~\cite[Appendix~C]{Moran2012}. The 157 characters from 12 
code blocks shown here are the characters proposed for ``strict'' IPA encoding, 
as discussed in Section~\ref{ipa-recommendations}.}

\begin{itemize}[itemsep=4pt]
	\item \textsc{Basic Latin }(27 characters) \newline 
	a b c d e f h i j k l m n o p q r s t u v w x y z~.~|
	\item \textsc{Latin-1 Supplement} (4 characters) \newline 
	æ ç ð ø
	\item \textsc{Latin Extended-A} (3 characters) \newline 
	ħ ŋ œ
	\item \textsc{Latin Extended-B} (4 characters) \newline 
	ǀ ǁ ǂ ǃ
	\item \textsc{IPA Extensions} (67 characters) \newline 
	ɐ ɑ ɒ ɓ ɔ ɕ ɖ ɗ ɘ ə ɛ ɜ ɞ ɟ ɠ ɡ ɢ ɣ ɤ ɥ ɦ ɧ ɨ ɪ ɬ ɭ ɮ ɯ ɰ ɱ ɲ ɳ ɴ \newline
	ɵ ɶ ɸ ɹ ɺ ɻ ɽ ɾ ʀ ʁ ʂ ʃ ʄ ʈ ʉ ʊ ʋ ʌ ʍ ʎ ʏ ʐ ʑ ʒ ʔ ʕ ʘ ʙ ʛ ʜ ʝ ʟ ʡ ʢ 
	\item \textsc{Spacing Modifier Letters} (17 characters) \newline
	\dia{02DE} \dia{02E1} \dia{02B7} \dia{02B2} \dia{02E0} \dia{02E4} 
	\dia{02B0} \dia{02BC} \dia{02D0} \dia{02D1} ˥ ˦ ˧ ˨ ˩ {\large ˈ ˌ}
	\item \textsc{Combining Diacritical Marks} (25 characters) \newline 
	\dia{033C} \dia{032A} \dia{033B} \dia{033A} \dia{031F} \dia{0320} \dia{031D} 
	\dia{031E} \dia{0318} \dia{0319} \dia{031C} \dia{0339} \dia{032C} \dia{0325} 
	\dia{0330} \dia{0324} \dia{0329} \dia{032F} \dia{0334} \dia{0303} \dia{0308} 
	\dia{033D} \dia{0306} \dia{031A} \ \dia{0361}◌
	\item \textsc{Greek and Coptic} (3 characters) \newline 
	β θ χ
    \item \textsc{General Punctuation} (2 characters) \newline 
    ‖ \charis{‿}
	\item \textsc{Superscripts and Subscripts} (1 character) \newline
	\dia{207F} 
	\item \textsc{Arrows} (4 characters) \newline 
	↑ ↓ ↗ ↘
	\item \textsc{Latin Extended-C} (1 character): \newline
	\charis{ⱱ}
\end{itemize}

@@

At the end of this chapter we have added a few longish tables summarizing all
159 different unicode codepoints that form the basis of strict-IPA encoding
(107 letters, 45 diacritics and 7 remaining symbols). Each of
these tables shows a typical glyph, and then lists the Unicode Codepoint,
Unicode Name and IPA description for each symbol.

@@

  \item \textsc{IPA letters} \newline
        The 107 different IPA letters as allowed in strict-IPA encoding are
        listed in Table~\ref{tab:ipa_letters} starting on
        page~\pageref{tab:ipa_letters}.
  \item \textsc{IPA diacritics} \newline The 45 different IPA diacritics and
        tone markers (both Unicode Modifier Letters and Combining Diacritical
        Marks) as allowed in strict-IPA encoding are listed in
        Table~\ref{tab:ipa_diacritics} starting on
        page~\pageref{tab:ipa_diacritics}.
  \item \textsc{IPA remainders} \newline
        The 7 remaining IPA symbols (boundary and stress markers) as allowed in
        strict-IPA encoding are listed in Table~\ref{tab:ipa_boundaries} on
        page~\pageref{tab:ipa_boundaries}. 
  \item \textsc{IPA additions} \newline
        The 10 proposed additions to the IPA are listed in
        Table~\ref{tab:ipa_additions} on page~\pageref{tab:ipa_additions}.

>>>>

As explained in Section~\ref{pitfall-blocks}, the Unicode Standard code space is
subdivided into character blocks, which generally encode characters from a
single script. However, as is illustrated by the IPA, characters that form a
single writing system may be dispersed across several different character
blocks. With its diverse collection of symbols from various scripts and
diacritics, the IPA is spread across 13 blocks in the Unicode
Standard:\footnote{This number of blocks depends on whether only IPA-sanctioned
symbols are counted or if the phonetic symbols commonly found in the literature
are also included, see~\cite[Appendix~C]{Moran2012}. The 159 characters from 13 
code blocks shown here are the characters proposed for ``strict'' IPA encoding, 
as discussed in Section~\ref{ipa-recommendations}.}

\begin{itemize}[itemsep=4pt]
	\item \textsc{Basic Latin }(27 characters) \newline 
	a b c d e f h i j k l m n o p q r s t u v w x y z~.~|
	\item \textsc{Latin-1 Supplement} (4 characters) \newline 
	æ ç ð ø
	\item \textsc{Latin Extended-A} (3 characters) \newline 
	ħ ŋ œ
	\item \textsc{Latin Extended-B} (4 characters) \newline 
	ǀ ǁ ǂ ǃ
	\item \textsc{IPA Extensions} (67 characters) \newline 
	ɐ ɑ ɒ ɓ ɔ ɕ ɖ ɗ ɘ ə ɛ ɜ ɞ ɟ ɠ ɡ ɢ ɣ ɤ ɥ ɦ ɧ ɨ ɪ ɬ ɭ ɮ ɯ ɰ ɱ ɲ ɳ ɴ \newline
	ɵ ɶ ɸ ɹ ɺ ɻ ɽ ɾ ʀ ʁ ʂ ʃ ʄ ʈ ʉ ʊ ʋ ʌ ʍ ʎ ʏ ʐ ʑ ʒ ʔ ʕ ʘ ʙ ʛ ʜ ʝ ʟ ʡ ʢ 
	\item \textsc{Spacing Modifier Letters} (17 characters) \newline
	\dia{02DE} \dia{02E1} \dia{02B7} \dia{02B2} \dia{02E0} \dia{02E4} 
	\dia{02B0} \dia{02BC} \dia{02D0} \dia{02D1} ˥ ˦ ˧ ˨ ˩ {\large ˈ ˌ}
	\item \textsc{Combining Diacritical Marks} (25 characters) \newline 
	\dia{033C} \dia{032A} \dia{033B} \dia{033A} \dia{031F} \dia{0320} \dia{031D} 
	\dia{031E} \dia{0318} \dia{0319} \dia{031C} \dia{0339} \dia{032C} \dia{0325} 
	\dia{0330} \dia{0324} \dia{0329} \dia{032F} \dia{0334} \dia{0303} \dia{0308} 
	\dia{033D} \dia{0306} \dia{031A} \ \dia{0361}◌
	\item \textsc{Greek and Coptic} (3 characters) \newline 
	β θ χ
    \item \textsc{General Punctuation} (2 characters) \newline 
    ‖ \charis{‿}
	\item \textsc{Superscripts and Subscripts} (1 character) \newline
	\dia{207F} 
	\item \textsc{Modifier Tone Letters} (2 characters) \newline
	ꜛ ꜜ
	\item \textsc{Arrows} (2 characters) \newline 
	↗ ↘
	\item \textsc{Latin Extended-C} (1 character): \newline
	\charis{ⱱ}
\end{itemize}

@@

At the end of this chapter we have added a few longish tables summarizing all
159 different unicode codepoints that form the basis of strict-IPA encoding
(107 letters, 41 diacritics and 11 remaining symbols). Each of
these tables shows a typical glyph, and then lists the Unicode Codepoint,
Unicode Name and IPA description for each symbol.

@@

   \item \textsc{IPA letters} \newline
        The 107 different IPA letters as allowed in strict-IPA encoding are
        listed in Table~\ref{tab:ipa_letters} starting on
        page~\pageref{tab:ipa_letters}.
  \item \textsc{IPA diacritics} \newline The 41 different IPA diacritics and
        tone markers (both Unicode Modifier Letters and Combining Diacritical
        Marks) as allowed in strict-IPA encoding are listed in
        Table~\ref{tab:ipa_diacritics} starting on
        page~\pageref{tab:ipa_diacritics}.
  \item \textsc{IPA remainders} \newline The 11 remaining IPA symbols
        (boundary, stress and intonation markers) as allowed in strict-IPA
        encoding are listed in Table~\ref{tab:ipa_boundaries} on
        page~\pageref{tab:ipa_boundaries}. 
  \item \textsc{IPA additions} \newline
        The 10 proposed additions to the IPA are listed in
        Table~\ref{tab:ipa_additions} on page~\pageref{tab:ipa_additions}.

### 2015-06-22 tables/unicode_boundaries.tex
updated handling of arrows, and changed codepoints
see wikipedia proposals: https://en.wikipedia.org/wiki/Upstep

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{5.6cm} L{4cm} }
{\large ˈ} & \unif{02C8} & \textsc{modifier letter vertical line} & primary stress \\
{\large ˌ} & \unif{02CC} & \textsc{modifier letter low vertical line} & secondary stress \\
{\large.} & \unif{002E} & \textsc{full stop} & syllable break \\
{|} & \unif{007C} & \textsc{vertical line} & minor group break (foot) \\
{‖} & \unif{2016} & \textsc{double vertical line} & major group break (intonation) \\
\charis{‿} & \unif{203F} & \textsc{undertie} & linking (absence of a break) \\
\end{xtabular}
\end{center}

>>>>

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{5.4cm} L{4cm} }
{\large ˈ} & \unif{02C8} & \textsc{modifier letter vertical line} & primary stress \\
{\large ˌ} & \unif{02CC} & \textsc{modifier letter low vertical line} & secondary stress \\
{ꜛ} & \unif{A71B} & \textsc{modifier letter raised up arrow} & upstep \\
{ꜜ} & \unif{A71C} & \textsc{modifier letter raised down arrow} & downstep \\
{↗} & \unif{2197} & \textsc{north east arrow} & global rise \\
{↘} & \unif{2198} & \textsc{south east arrow} & global fall \\
 & \unif{0020} & \textsc{space} & word break \\
{\large.} & \unif{002E} & \textsc{full stop} & syllable break \\
{|} & \unif{007C} & \textsc{vertical line} & minor group break (foot) \\
{‖} & \unif{2016} & \textsc{double vertical line} & major group break (intonation) \\
\charis{‿} & \unif{203F} & \textsc{undertie} & linking (absence of a break) \\
\end{xtabular}
\end{center}

### 2015-06-22 tables/unicode_diacritics.tex
adding tilde overlay and tie bar to strict IPA
updated handling of arrows, and changed codepoints
see wikipedia proposals: https://en.wikipedia.org/wiki/Upstep

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{6.6cm} L{3cm} }
\dia{033C} & \unif{033C} & \textsc{combining seagull below} & linguolabial \\
\dia{032A} & \unif{032A} & \textsc{combining bridge below} & dental \\
\dia{033B} & \unif{033B} & \textsc{combining square below} & laminal \\
\dia{033A} & \unif{033A} & \textsc{combining inverted bridge below} & apical \\
\dia{031F} & \unif{031F} & \textsc{combining plus sign below} & advanced \\
\dia{0320} & \unif{0320} & \textsc{combining minus sign below} & retracted \\
\dia{031D} & \unif{031D} & \textsc{combining up tack below} & raised \\
\dia{031E} & \unif{031E} & \textsc{combining down tack below} & lowered \\
\dia{0318} & \unif{0318} & \textsc{combining left tack below} & advanced tongue root \\
\dia{0319} & \unif{0319} & \textsc{combining right tack below} & retracted tongue root \\
\dia{031C} & \unif{031C} & \textsc{combining left half ring below} & less rounded \\
\dia{0339} & \unif{0339} & \textsc{combining right half ring below} & more rounded \\
\dia{032C} & \unif{032C} & \textsc{combining caron below} & voiced \\
\dia{0325} & \unif{0325} & \textsc{combining ring below} & voiceless \\
\dia{0330} & \unif{0330} & \textsc{combining tilde below} & creaky voiced \\
\dia{0324} & \unif{0324} & \textsc{combining diaeresis below} & breathy voiced \\
\dia{0329} & \unif{0329} & \textsc{combining vertical line below} & syllabic \\
\dia{032F} & \unif{032F} & \textsc{combining inverted breve below} & non-syllabic \\
\dia{0303} & \unif{0303} & \textsc{combining tilde} & nasalized \\
\dia{0308} & \unif{0308} & \textsc{combining diaeresis} & centralized \\
\dia{033D} & \unif{033D} & \textsc{combining x above} & mid-centralized \\
\dia{0306} & \unif{0306} & \textsc{combining breve} & extra-short \\
\dia{031A} & \unif{031A} & \textsc{combining left angle above} & no audible release \\
\dia{02DE} & \unif{02DE} & \textsc{modifier letter rhotic hook} & rhotacized \\
\dia{02E1} & \unif{02E1} & \textsc{modifier letter small l} & lateral release \\
\dia{207F} & \unif{207F} & \textsc{superscript latin small letter n} & nasal release \\
\dia{02B7} & \unif{02B7} & \textsc{modifier letter small w} & labialized \\
\dia{02B2} & \unif{02B2} & \textsc{modifier letter small j} & palatalized \\
\dia{02E0} & \unif{02E0} & \textsc{modifier letter small gamma} & velarized \\
\dia{02E4} & \unif{02E4} & \textsc{modifier letter small reversed glottal stop} & pharyngealized \\
\dia{02B0} & \unif{02B0} & \textsc{modifier letter small h} & aspirated \\
\dia{02BC} & \unif{02BC} & \textsc{modifier letter apostrophe} & ejective \\
\dia{02D0} & \unif{02D0} & \textsc{modifier letter triangular colon} & long \\
\dia{02D1} & \unif{02D1} & \textsc{modifier letter half triangular colon} & half-long \\{˥} & \unif{02E5} & \textsc{modifier letter extra-high tone bar} & extra high tone \\
{˦} & \unif{02E6} & \textsc{modifier letter high tone bar} & high tone \\
{˧} & \unif{02E7} & \textsc{modifier letter mid tone bar} & mid tone \\
{˨} & \unif{02E8} & \textsc{modifier letter low tone bar} & low tone \\
{˩} & \unif{02E9} & \textsc{modifier letter extra-low tone bar} & extra low tone \\
{↑} & \unif{2191} & \textsc{upwards arrow} & upstep \\
{↓} & \unif{2193} & \textsc{downwards arrow} & downstep \\
{↗} & \unif{2197} & \textsc{north east arrow} & global rise \\
{↘} & \unif{2198} & \textsc{south east arrow} & global fall \\
\end{xtabular}
\end{center}

>>>>

\begin{center}
\begin{xtabular}{ l L{1.1cm} L{6.2cm} L{3.6cm} }
\dia{0334} & \unif{0334} & \textsc{combining tilde overlay} & velarized or pharyngealized \\
\dia{033C} & \unif{033C} & \textsc{combining seagull below} & linguolabial \\
\dia{032A} & \unif{032A} & \textsc{combining bridge below} & dental \\
\dia{033B} & \unif{033B} & \textsc{combining square below} & laminal \\
\dia{033A} & \unif{033A} & \textsc{combining inverted bridge below} & apical \\
\dia{031F} & \unif{031F} & \textsc{combining plus sign below} & advanced \\
\dia{0320} & \unif{0320} & \textsc{combining minus sign below} & retracted \\
\dia{031D} & \unif{031D} & \textsc{combining up tack below} & raised \\
\dia{031E} & \unif{031E} & \textsc{combining down tack below} & lowered \\
\dia{0318} & \unif{0318} & \textsc{combining left tack below} & advanced tongue root \\
\dia{0319} & \unif{0319} & \textsc{combining right tack below} & retracted tongue root \\
\dia{031C} & \unif{031C} & \textsc{combining left half ring below} & less rounded \\
\dia{0339} & \unif{0339} & \textsc{combining right half ring below} & more rounded \\
\dia{032C} & \unif{032C} & \textsc{combining caron below} & voiced \\
\dia{0325} & \unif{0325} & \textsc{combining ring below} & voiceless \\
\dia{0330} & \unif{0330} & \textsc{combining tilde below} & creaky voiced \\
\dia{0324} & \unif{0324} & \textsc{combining diaeresis below} & breathy voiced \\
\dia{0329} & \unif{0329} & \textsc{combining vertical line below} & syllabic \\
\dia{032F} & \unif{032F} & \textsc{combining inverted breve below} & non-syllabic \\
\dia{0303} & \unif{0303} & \textsc{combining tilde} & nasalized \\
\dia{0308} & \unif{0308} & \textsc{combining diaeresis} & centralized \\
\dia{033D} & \unif{033D} & \textsc{combining x above} & mid-centralized \\
\dia{0306} & \unif{0306} & \textsc{combining breve} & extra-short \\
\dia{031A} & \unif{031A} & \textsc{combining left angle above} & no audible release \\
\dia{02DE} & \unif{02DE} & \textsc{modifier letter rhotic hook} & rhotacized \\
\dia{02E1} & \unif{02E1} & \textsc{modifier letter small l} & lateral release \\
\dia{207F} & \unif{207F} & \textsc{superscript latin small letter n} & nasal release \\
\dia{02B7} & \unif{02B7} & \textsc{modifier letter small w} & labialized \\
\dia{02B2} & \unif{02B2} & \textsc{modifier letter small j} & palatalized \\
\dia{02E0} & \unif{02E0} & \textsc{modifier letter small gamma} & velarized \\
\dia{02E4} & \unif{02E4} & \textsc{modifier letter small reversed glottal stop} & pharyngealized \\
\dia{02B0} & \unif{02B0} & \textsc{modifier letter small h} & aspirated \\
\dia{02BC} & \unif{02BC} & \textsc{modifier letter apostrophe} & ejective \\
\dia{02D0} & \unif{02D0} & \textsc{modifier letter triangular colon} & long \\
\dia{02D1} & \unif{02D1} & \textsc{modifier letter half triangular colon} & half-long \\{˥} & \unif{02E5} & \textsc{modifier letter extra-high tone bar} & extra high tone \\
{˦} & \unif{02E6} & \textsc{modifier letter high tone bar} & high tone \\
{˧} & \unif{02E7} & \textsc{modifier letter mid tone bar} & mid tone \\
{˨} & \unif{02E8} & \textsc{modifier letter low tone bar} & low tone \\
{˩} & \unif{02E9} & \textsc{modifier letter extra-low tone bar} & extra low tone \\
\dia{0361}{\large\fontspec{CharisSIL}◌} & \unif{0361} & \textsc{combining double inverted breve} & tie bar \\
\end{xtabular}
\end{center}
